\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-12-02 Documentation/TearingModeTheory.tex
Finished first part of tearing mode write up

\begin{document}
\title{\bf Tearing Mode Dynamics in Tokamak Plasmas}
\author{R.~Fitzpatrick\,\footnote{rfitzp@utexas.edu}\\
Institute for Fusion Studies,\\  Department of Physics,\\  University of Texas at Austin,\\  Austin TX 78712, USA}
\date{}
\maketitle

@@

\subsection{Cylindrical Tearing Mode Equation}
To lowest order, the tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Wesson 1978; Fitzpatrick 1993),
\begin{equation}
\frac{\partial^2\psi}{\partial r^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}=  0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}

@@

Note that Equation~(\ref{e100}) is singular at the so-called {\em resonant}\/ magnetic flux-surface, radius $r=r_s$, at which 
\begin{equation}\label{e104}
q(r_s)= \frac{m}{n}.
\end{equation}
At the resonant surface, ${\bf k}\cdot{\bf B}=0$, where ${\bf k} = (0,\,m/r,\,-n/R_0)$ is the wavevector of the tearing perturbation. 
The value of $\psi(r,t)$ at the resonant surface---namely, ${\mit\Psi}_s(t)$---is known as the {\em reconnected magnetic flux}\/ (Fitzpatrick 1993). (See Section~\ref{dis1}.)
Note that ${\mit\Psi}_s$ is, in general, a complex quantity. 
Let $\rho=(r-r_s)/r_s$. 
The solution of Equation~(\ref{e100}) in the vicinity of the resonant surface is
\begin{equation}\label{e105}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s+}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho>0$, and
\begin{equation}\label{e106}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s-}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho<0$. Here, 
\begin{align}
\alpha_s& =- \left(\frac{r\,q\,J_z'}{s}\right)_{r=r_s},\\[0.5ex]
s(r) &= \frac{r\,q'}{q}.\label{shear}
\end{align}
Moreover,  the real parameters ${\mit\Delta}_{s+}$ and  ${\mit\Delta}_{s-}$ are fully determined by Equation~(\ref{e100})
and the boundary conditions (\ref{e101})--(\ref{e103}). Note that, in general, $\psi$ is continuous across the resonant surface (in accordance with Maxwell's equations),
whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial \psi/\partial r$ implies the presence of a radially-thin, helical,  current sheet  at the resonant surface. This current sheet can only be resolved by retaining more terms in the perturbed plasma equation of motion, (\ref{e96}). (See Section~\ref{linear}.)

@@

\subsection{Resonant Layer Physics}\label{sres}
By analogy with Equation~(\ref{e133}), we can write
\begin{equation}\label{e134}
{\mit\Delta\Psi}_s = \tau_s\left(\frac{d{\mit\Psi}_s}{dt} + {\rm i}\,\omega_s\,{\mit\Psi}_s\right).
\end{equation}
Here, $\tau_s$ is the {\em reconnection time}\/ (i.e., the typical timescale on which magnetic reconnection takes place in the
resonant layer surrounding the resonant surface), whereas $\omega_s$ is the so-called {\em natural frequency}. 
The natural frequency is the phase velocity of the tearing mode when the wall is perfectly conducting. The
natural frequency is non-zero because the reconnected flux in the resonant layer is convected by the local plasma
flow (Fitzpatrick 1993). (See Section~\ref{s6.2}.)

@@

In general, $\psi$ is continuous across the field-coil (in accordance with Maxwell's equations), whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial\psi/\partial r$ is caused by the helical currents flowing in the coil. The complex quantity ${\mit\Psi}_c(t)$ determines the amplitude and
phase of the perturbed magnetic flux at the coil. The complex quantity
\begin{equation}\label{e118a}
{\mit\Delta\Psi}_c = \left[r\,\frac{\partial\psi}{\partial r}\right]_{r_{c-}}^{r_{c+}}
\end{equation}
parameterizes the amplitude and phase of the helical  current sheet flowing in the coil.
It follows from Equations~(\ref{e93a}), (\ref{e94}), and (\ref{e135})  that
\begin{equation}\label{e145}
{\mit\Delta\Psi}_c = -\mu_0\,I_c.
\end{equation}

@@

\subsection{Electromagnetic Torques}\label{storque}
The flux-surface integrated poloidal and toroidal electromagnetic torque densities acting on the plasma
can be written
\begin{align}\label{e152}
T_\theta(r) &= \left\langle r\,{\bf j}\times {\bf B}\cdot {\bf e}_\theta\right\rangle\\[0.5ex]
T_\varphi(r) &= \left\langle R_0\,{\bf j}\times {\bf B}\cdot {\bf e}_z\right\rangle,
\end{align}
respectively, 
where
\begin{equation}
\left\langle\cdots\right\rangle \equiv \oint\oint r\,R_0\,(\cdots)\,d\theta\,d\varphi
\end{equation}
is a flux-surface integration  operator. 
 However, according to Equation~(\ref{e80}),
\begin{equation}\label{e150}
{\bf j}\times {\bf B} \simeq \nabla p.
\end{equation}
Given that the scalar pressure is a single-valued function of $\theta$ and $\varphi$, it immediately follows
that $T_\theta=T_\varphi = 0$ throughout the plasma (Fitzpatrick 1993). The only exception to this rule occurs in the immediate vicinity of the
resonant surface, where Equation~(\ref{e150}) breaks down. It follows that we can write
\begin{align}\label{e156a}
T_\theta(r) &= T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
T_\varphi(r) &= T_{\varphi\,s}\,\delta(r-r_s),
\end{align}
where
\begin{align}\label{e153}
T_{\theta\,s} &=  \frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0\,r^{\,2}\,(\delta j_z\,\delta B_r^{\,\ast} + \delta j_z^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\varphi,\\[0.5ex]
T_{\varphi\,s} &=  -\frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0^{\,2}\,r\,(\delta j_\theta\,\delta B_r^{\,\ast} + \delta j_\theta^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\varphi
\end{align}
are the net poloidal and toroidal torques, respectively, acting at the resonant surface. 
Note that the zeroth-order (in perturbed quantities) torques are zero because $B_r=0$. Furthermore, the linear (in perturbed quantities) torques 
average to zero over the flux-surface. Hence, the largest non-zero torques  are {\em quadratic}\/ in perturbed quantities. 
According to Equation~(\ref{e82}), 
\begin{equation}
\nabla \cdot \delta {\bf j} = {\rm i}\left(\frac{m}{r}\,\delta j_\theta - \frac{n}{R_0}\,\delta j_z\right) = 0,
\end{equation}
where $\delta{\bf j}$ is the density of the current sheet flowing at the resonant surface, and we have made use
of the fact that $\delta j_r$ is negligible in a radially-thin current sheet (Fitzpatrick 1993). 
 It follows from the previous three equations that
\begin{equation}\label{e156}
T_{\varphi\,s} = - \frac{n}{m}\,T_{\theta\,s}.
\end{equation}
Finally, Equations~(\ref{e93}), (\ref{e94}), (\ref{e153}), and (\ref{e156}) imply that
\begin{align}\label{e157}
T_{\theta\,s} &= -\frac{2\pi^2\,R_0\,m}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast}),\\[0.5ex]
T_{\varphi\,s} &= \frac{2\pi^2\,R_0\,n}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast})\label{e158}
\end{align}
(Fitzpatrick 1993).

@@

Taking  the scalar product of $r\,{\bf e}_\theta$ with the plasma equation of motion, (\ref{e175}), and integrating around magnetic
flux-surfaces, we obtain 
\begin{equation}\label{e176}
\left\langle\rho\,\frac{\partial (r\,V_{\theta\,i})}{\partial t} \right\rangle 
+\left\langle \rho\left[V_{r\,i}\,\frac{\partial (r\,V_{\theta\,i})}{\partial r} + V_{z\,i}\,\frac{\partial(r\, V_{\theta\,i})}{\partial z}
\right]\right\rangle + [\zeta_i^{-1}]\,\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle
= [\delta_i^{-2}]\,T_{\theta\,s}\,\delta(r-r_s),
\end{equation}
where use has been made of Equations~(\ref{e152}) and (\ref{e156a}). Note that the dominant $\nabla p$ term
has averaged to zero. Moreover, because $T_{\theta\,s}$ is quadratic in perturbed quantities (see Section~\ref{storque}), it is plausible that the right-hand side of the previous equation is no longer a dominant term [i.e., it is not ${\cal O}(\delta_i^{-2})$ larger than the
other terms in the equation]. To be more exact, if $|\delta {\bf B}|/|{\bf B}|\sim \delta_i$ then the right-hand side ceases to be a dominant term in the equation.

The largest contribution to $\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle$
comes from the ion parallel viscosity tensor acting on the zeroth-order (in perturbed quantities) ion flow. (See Section~\ref{sbrag}.)  However, this contribution averages to zero in a cylindrical plasma, in which $B_z$
is constant on magnetic flux-surfaces. Of course, in reality, in a large aspect-ratio tokamak plasma, 
$B_z$ varies slightly around magnetic flux-surfaces, being larger on the inboard side of the torus than on the outboard side (Wesson 2011). This variation gives rise to a residual contribution to $\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle$.
It is plausible that this contribution is not a dominant term in Equation~(\ref{e176})  [i.e., it is not ${\cal O}(\zeta_i^{-1})$ larger than the
other terms in the equation]. To be more exact, if the residual contribution is ${\cal O}(\zeta_i)$ times smaller than the non-averaged  contribution then  $\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle$ ceases to be a dominant term in the equation.

According to the previous discussion, it is plausible that all terms appearing in Equation~(\ref{e176}) are of approximately equal
magnitude when the inertial and viscous terms are evaluated with the zeroth-order ion flow, which can be written
\begin{equation}
{\bf V}_i = r\,{\mit\Omega}_\theta(r,t)\,{\bf e}_\theta + R_0\,{\mit\Omega}_\varphi(r,t)\,{\bf e}_z.
\end{equation}
Here, ${\mit\Omega}_\theta(r,t)$ and ${\mit\Omega}_\varphi(r,t)$ are the ion poloidal and toroidal angular velocity
profiles, respectively. When $\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle$ is evaluated with  the ion parallel viscosity tensor, and the zeroth-order
ion flow, it gives rise to a term that acts to relax the ion poloidal angular velocity profile toward a so-called {\em neoclassical
profile}\/ determined by ion density and temperature gradients (Stix 1973; Hirshman \& Sigmar 1981). In principle, the
contribution of the ion perpendicular viscosity tensor to $\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle$ 
is ${\cal O}(\zeta_i^{\,2})$ smaller than the contribution of the ion parallel viscosity tensor. (See Section~\ref{sbrag}.) In practice, the magnitude of the ion perpendicular
viscosity tensor is greatly enhanced above the value predicted by the Braginskii equations by the action of small-scale
plasma turbulence (Wesson 2011). Hence, it is reasonable to include the contribution of the enhanced perpendicular
viscosity tensor to $\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle$. Our model form for $\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle$ is written
\begin{equation}\label{e178}
\langle r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\rangle = 4\pi^2\,R_0\left[ \frac{\rho}{\tau_\theta}\,r^{\,3}\,({\mit\Omega}_\theta-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^{\,3}\,\frac{\partial{\mit\Omega}_\theta}{\partial r}\right)\right].
\end{equation}
Here, $\rho(r)$ is the equilibrium plasma mass density profile, $\tau_\theta(r)$ is the {\em poloidal flow-damping time}\/ [which
is generally of order $(a/q\,R_0)^2\,\tau_i$], ${\mit\Omega}_{\theta\,{\rm nc}}(r)$ is the neoclassical poloidal angular velocity profile, and
$\mu_\perp(r)$ is a phenomenological perpendicular viscosity due to plasma turbulence.

@@

\subsection{No-Slip Constraint}\label{sns}
The reconnected magnetic flux at the resonant surface is assumed to convected by the local plasma flow. (See Section~\ref{sres}.) Hence, changes in the
plasma flow at the resonant surface induced by the electromagnetic torques that develop in the vicinity of this
surface will modify the convection velocity. Consequently, the natural frequency of the reconnected flux can be written
\begin{equation}\label{e200}
 \omega_s(t) = \omega_{s\,0} + ({\bf k}\cdot {\mit\Delta}{\bf V}_i)_{r=r_s}=
 \omega_{s\,0} + m\,{\mit\Delta\Omega}_\theta(r_s,t) - n\,{\mit\Delta\Omega}_\varphi(r_s,t),
\end{equation}
where $\omega_{s\,0}$ is the unperturbed natural frequency. The previous expression is known as the {\em no-slip constraint}\/ (Fitzpatrick 1993), and has been verified experimentally (Vahala, et al.\ 1980). 
It follows from Equations~(\ref{e190})
and (\ref{e191}) that
\begin{equation}\label{e201}
\omega_s(t) = \omega_{s\,0} - \sum_{p=1,\infty}\left[\alpha_p(t)+\beta_p(t)\right].
\end{equation}
(See Section~\ref{stb}.)

\subsection{Final Equations}
Equations~(\ref{e123}), (\ref{e133}), (\ref{e134}),  (\ref{e145})--(\ref{e151}), (\ref{e198}), (\ref{e199}), and (\ref{e201})
can be combined to give the following complete set of equations that determine the time evolution of the reconnected 
magnetic flux at the resonant surface:
\begin{align}\label{e202}
{\mit\Delta\hat{\Psi}}_s\equiv \tau_s\left(\frac{d\hat{\mit\Psi}_s}{dt} + {\rm i}\,\omega_s\,\hat{\mit\Psi}_s\right) &= E_{ss}\,\hat{\mit\Psi}_s + E_{sw}\,\hat{\mit\Psi}_w,\\[0.5ex]
\tau_w\,\frac{d\hat{\mit\Psi}_w}{dt} &= E_{sw}\,\hat{\mit\Psi}_s +\tilde{E}_{ww}\,\hat{\mit\Psi}_w
+\hat{I}_c,\label{e203}\\[0.5ex]
\omega_s(t)& = \omega_{s\,0} - \sum_{p=1,\infty}\left[\alpha_p(t)+\beta_p(t)\right],\\[0.5ex]
(1+2\,q_s^{\,2})\,\frac{d\alpha_p}{dt} + \left(\frac{1}{\tau_{\theta\,s}}+\frac{j_{1\,p}^{\,2}}{\tau_{M}}\right)\alpha_p
&= \frac{m^{2}\,[y_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,\epsilon_a^{\,2}\,[J_2(j_{1\,p})]^{\,2}}
\,{\rm Im}({\mit\Delta\hat{\Psi}}_s\,\hat{\mit\Psi}_s^{\,\ast}),\label{e205}\\[0.5ex]
\frac{d\beta_p}{dt} + \frac{j_{0\,p}^{\,2}}{\tau_{M}}\,\beta_p
&= \frac{n^{2}\,[z_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,[J_1(j_{0\,p})]^{\,2}}\,
{\rm Im}({\mit\Delta\hat{\Psi}}_s\,\hat{\mit\Psi}_s^{\,\ast}).\label{e206}
\end{align}
Here, 
\begin{align}
\hat{I}_c &= \left(\frac{r_w}{r_c}\right)^m\frac{\mu_0\,I_c}{R_0\,B_z},\\[0.5ex]
\tilde{E}_{ww} &= E_{ww} + \frac{2m\,(r_w/r_c)^{2m}}{1-(r_w/r_c)^{2m}}.\label{e208}
\end{align}

>>>>

\begin{document}
\title{\bf Tearing Mode Dynamics in Tokamak Plasmas}
\author{R.~Fitzpatrick\,\footnote{rfitzp@utexas.edu}\\[0.5ex]
Institute for Fusion Studies,\\  Department of Physics,\\  University of Texas at Austin,\\  Austin TX 78712, USA}
\date{}
\maketitle

@@

\subsection{Cylindrical Tearing Mode Equation}
To lowest order, the tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Wesson 1978; Fitzpatrick 1993),
\begin{equation}\label{e97}
\frac{\partial^2\psi}{\partial r^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}=  0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}

@@

Note that Equation~(\ref{e100}) is singular at the so-called {\em resonant}\/ magnetic flux-surface, radius $r=r_s$, at which 
\begin{equation}\label{e104}
q(r_s)= \frac{m}{n}.
\end{equation}
At the resonant surface, ${\bf k}\cdot{\bf B}=0$, where ${\bf k} = (0,\,m/r,\,-n/R_0)$ is the wavevector of the tearing perturbation. 
The value of $\psi(r,t)$ at the resonant surface---namely, ${\mit\Psi}_s(t)$---is known as the {\em reconnected magnetic flux}\/ (Fitzpatrick 1993). (See Section~\ref{dis1}.)
Note that ${\mit\Psi}_s$ is, in general, a complex quantity. 
Let $\rho=(r-r_s)/r_s$. 
The solution of Equation~(\ref{e100}) in the vicinity of the resonant surface is
\begin{equation}\label{e105}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s+}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho>0$, and
\begin{equation}\label{e106}
\hat{\psi}_s(\rho) = 1 + {\mit\Delta}_{s-}\,\rho + \alpha_s\,\rho\,\ln|\rho| + {\cal O}(\rho^{\,2})
\end{equation}
for $\rho<0$. Here, 
\begin{align}\label{e108}
\alpha_s& =- \left(\frac{r\,q\,J_z'}{s}\right)_{r=r_s},\\[0.5ex]
s(r) &= \frac{r\,q'}{q}.\label{shear}
\end{align}
Moreover,  the real parameters ${\mit\Delta}_{s+}$ and  ${\mit\Delta}_{s-}$ are fully determined by Equation~(\ref{e100})
and the boundary conditions (\ref{e101})--(\ref{e103}). Note that, in general, $\psi$ is continuous across the resonant surface (in accordance with Maxwell's equations),
whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial \psi/\partial r$ implies the presence of a radially-thin, helical,  current sheet  at the resonant surface. This current sheet can only be resolved by retaining more terms in the perturbed plasma equation of motion, (\ref{e96}). (See Section~\ref{linear}.)

@@

\subsection{Resonant Layer Physics}\label{sres}
By analogy with Equation~(\ref{e133}), we can write
\begin{equation}\label{e134}
{\mit\Delta\Psi}_s = \tau_s\left(\frac{d{\mit\Psi}_s}{dt} + {\rm i}\,\omega\,{\mit\Psi}_s\right).
\end{equation}
Here, $\tau_s$ is the {\em reconnection time}\/ (i.e., the typical timescale on which magnetic reconnection takes place in the
resonant layer surrounding the resonant surface), whereas $\omega$ is the phase velocity 
of the tearing mode when the wall is perfectly conducting. The
phase velocity  is non-zero because the reconnected flux in the resonant layer is convected by the local plasma
flow (Fitzpatrick 1993). (See Sections~\ref{s6.2} and \ref{sphase1}.)

@@

In general, $\psi$ is continuous across the field-coil (in accordance with Maxwell's equations), whereas $\partial\psi/\partial r$ is discontinuous. The discontinuity in $\partial\psi/\partial r$ is caused by the helical current flowing in the coil. The complex quantity ${\mit\Psi}_c(t)$ determines the amplitude and
phase of the perturbed magnetic flux at the coil. The complex quantity
\begin{equation}\label{e118a}
{\mit\Delta\Psi}_c = \left[r\,\frac{\partial\psi}{\partial r}\right]_{r_{c-}}^{r_{c+}}
\end{equation}
parameterizes the amplitude and phase of the helical  current sheet flowing in the coil.
It follows from Equations~(\ref{e93a}), (\ref{e94}), and (\ref{e135})  that
\begin{equation}\label{e145}
{\mit\Delta\Psi}_c = -\mu_0\,I_c.
\end{equation}

@@

\subsection{Electromagnetic Torques}\label{storque}
The flux-surface integrated poloidal and toroidal electromagnetic torque densities acting on the plasma
can be written
\begin{align}\label{e152}
T_\theta(r) &= \left\{ r\,{\bf j}\times {\bf B}\cdot {\bf e}_\theta\right\}\\[0.5ex]
T_\varphi(r) &= \left\{ R_0\,{\bf j}\times {\bf B}\cdot {\bf e}_z\right\},
\end{align}
respectively, 
where
\begin{equation}
\left\{\cdots\right\} \equiv \oint\oint r\,R_0\,\cdots\,d\theta\,d\varphi
\end{equation}
is a flux-surface integration  operator. 
 However, according to Equation~(\ref{e80}),
\begin{equation}\label{e150}
{\bf j}\times {\bf B} \simeq \nabla p.
\end{equation}
Given that the scalar pressure is a single-valued function of $\theta$ and $\varphi$, it immediately follows
that $T_\theta=T_\varphi = 0$ throughout the plasma (Fitzpatrick 1993). The only exception to this rule occurs in the immediate vicinity of the
resonant surface, where Equation~(\ref{e150}) breaks down. It follows that we can write
\begin{align}\label{e156a}
T_\theta(r) &= T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
T_\varphi(r) &= T_{\varphi\,s}\,\delta(r-r_s),
\end{align}
where
\begin{align}\label{e153}
T_{\theta\,s} &=  \frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0\,r^{\,2}\,(\delta j_z\,\delta B_r^{\,\ast} + \delta j_z^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\varphi,\\[0.5ex]
T_{\varphi\,s} &=  -\frac{1}{4}\int_{r_{s-}}^{r_{s+}}\oint\oint R_0^{\,2}\,r\,(\delta j_\theta\,\delta B_r^{\,\ast} + \delta j_\theta^{\,\ast}\,\delta B_r)\,dr\,d\theta\,d\varphi
\end{align}
are the net poloidal and toroidal torques, respectively, acting at the resonant surface. 
Note that the zeroth-order (in perturbed quantities) torques are zero because $B_r=0$. Furthermore, the linear (in perturbed quantities) torques 
average to zero over the flux-surface. Hence, the largest non-zero torques  are {\em quadratic}\/ in perturbed quantities. 
According to Equation~(\ref{e82}), 
\begin{equation}
\nabla \cdot \delta {\bf j} = {\rm i}\left(\frac{m}{r}\,\delta j_\theta - \frac{n}{R_0}\,\delta j_z\right) = 0,
\end{equation}
where $\delta{\bf j}$ is the density of the current sheet flowing at the resonant surface, and we have made use
of the fact that $\delta j_r$ is negligible in a radially-thin current sheet (Fitzpatrick 1993). 
 It follows from the previous three equations that
\begin{equation}\label{e156}
T_{\varphi\,s} = - \frac{n}{m}\,T_{\theta\,s}.
\end{equation}
Finally, Equations~(\ref{e93}), (\ref{e94}), (\ref{e153}), and (\ref{e156}) imply that
\begin{align}\label{e157}
T_{\theta\,s} &= -\frac{2\pi^2\,R_0\,m}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast}),\\[0.5ex]
T_{\varphi\,s} &= \frac{2\pi^2\,R_0\,n}{\mu_0}\,{\rm Im}({\mit\Delta\Psi}_s\,{\mit\Psi}_s^{\,\ast})\label{e158}
\end{align}
(Fitzpatrick 1993).

@@

Taking  the scalar product of $r\,{\bf e}_\theta$ with the plasma equation of motion, (\ref{e175}), and integrating around magnetic
flux-surfaces, we obtain 
\begin{equation}\label{e176}
\left\{\rho\,\frac{\partial (r\,V_{\theta\,i})}{\partial t} \right\} 
+\left\{ \rho\left[V_{r\,i}\,\frac{\partial (r\,V_{\theta\,i})}{\partial r} + V_{z\,i}\,\frac{\partial(r\, V_{\theta\,i})}{\partial z}
\right]\right\} + [\zeta_i^{-1}]\,\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}
= [\delta_i^{-2}]\,T_{\theta\,s}\,\delta(r-r_s),
\end{equation}
where use has been made of Equations~(\ref{e152}) and (\ref{e156a}). Note that the dominant $\nabla p$ term
has averaged to zero. Moreover, because $T_{\theta\,s}$ is quadratic in perturbed quantities (see Section~\ref{storque}), it is plausible that the right-hand side of the previous equation is no longer a dominant term [i.e., it is not ${\cal O}(\delta_i^{-2})$ larger than the
other terms in the equation]. To be more exact, if $|\delta {\bf B}|/|{\bf B}|\sim \delta_i$ then the right-hand side ceases to be a dominant term in the equation.

The largest contribution to $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$
comes from the ion parallel viscosity tensor acting on the zeroth-order (in perturbed quantities) ion flow. (See Section~\ref{sbrag}.)  However, this contribution averages to zero in a cylindrical plasma, in which $B_z$
is constant on magnetic flux-surfaces. Of course, in reality, in a large aspect-ratio tokamak plasma, 
$B_z$ varies slightly around magnetic flux-surfaces, being larger on the inboard side of the torus than on the outboard side (Wesson 2011). This variation gives rise to a residual contribution to $\{r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$.
It is plausible that this contribution is not a dominant term in Equation~(\ref{e176})  [i.e., it is not ${\cal O}(\zeta_i^{-1})$ larger than the
other terms in the equation]. To be more exact, if the residual contribution is ${\cal O}(\zeta_i)$ times smaller than the non-averaged  contribution then  $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ ceases to be a dominant term in the equation.

According to the previous discussion, it is plausible that all terms appearing in Equation~(\ref{e176}) are of approximately equal
magnitude when the inertial and viscous terms are evaluated with the zeroth-order ion flow, which can be written
\begin{equation}
{\bf V}_i = r\,{\mit\Omega}_\theta(r,t)\,{\bf e}_\theta + R_0\,{\mit\Omega}_\varphi(r,t)\,{\bf e}_z.
\end{equation}
Here, ${\mit\Omega}_\theta(r,t)$ and ${\mit\Omega}_\varphi(r,t)$ are the ion poloidal and toroidal angular velocity
profiles, respectively. When $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ is evaluated with  the ion parallel viscosity tensor, and the zeroth-order
ion flow, it gives rise to a term that acts to relax the ion poloidal angular velocity profile toward a so-called {\em neoclassical
profile}\/ determined by ion density and temperature gradients (Stix 1973; Hirshman \& Sigmar 1981). In principle, the
contribution of the ion perpendicular viscosity tensor to $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ 
is ${\cal O}(\zeta_i^{\,2})$ smaller than the contribution of the ion parallel viscosity tensor. (See Section~\ref{sbrag}.) In practice, the magnitude of the ion perpendicular
viscosity tensor is greatly enhanced above the value predicted by the Braginskii equations by the action of small-scale
plasma turbulence (Wesson 2011). Hence, it is reasonable to include the contribution of the enhanced perpendicular
viscosity tensor to $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$. Our model form for $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ is written
\begin{equation}\label{e178}
\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\} = 4\pi^2\,R_0\left[ \frac{\rho}{\tau_\theta}\,r^{\,3}\,({\mit\Omega}_\theta-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^{\,3}\,\frac{\partial{\mit\Omega}_\theta}{\partial r}\right)\right].
\end{equation}
Here, $\rho(r)$ is the equilibrium plasma mass density profile, $\tau_\theta(r)$ is the {\em poloidal flow-damping time}\/ [which
is generally of order $(a/q\,R_0)^2\,\tau_i$], ${\mit\Omega}_{\theta\,{\rm nc}}(r)$ is the neoclassical poloidal angular velocity profile, and
$\mu_\perp(r)$ is a phenomenological perpendicular viscosity due to plasma turbulence.

@@

\subsection{No-Slip Constraint}\label{sns}
The reconnected magnetic flux at the resonant surface is assumed to convected by the local plasma flow. (See Section~\ref{sres}.) Hence, changes in the
plasma flow at the resonant surface induced by the electromagnetic torques that develop in the vicinity of this
surface will modify the convection velocity. Consequently, the phase velocity of the tearing mode can be written
\begin{equation}\label{e200}
 \omega(t) = \omega_{0} + ({\bf k}\cdot {\mit\Delta}{\bf V}_i)_{r=r_s}=
 \omega_{0} + m\,{\mit\Delta\Omega}_\theta(r_s,t) - n\,{\mit\Delta\Omega}_\varphi(r_s,t),
\end{equation}
where $\omega_{0}$ is the unperturbed phase velocity. The previous expression is known as the {\em no-slip constraint}\/ (Fitzpatrick 1993), and has been verified experimentally (Vahala, et al.\ 1980). 
It follows from Equations~(\ref{e190})
and (\ref{e191}) that
\begin{equation}\label{e201}
\omega(t) = \omega_{0} - \sum_{p=1,\infty}\left[\alpha_p(t)+\beta_p(t)\right].
\end{equation}
(See Section~\ref{stb}.)

\subsection{Final Equations}
Equations~(\ref{e123}), (\ref{e133}), (\ref{e134}),  (\ref{e145})--(\ref{e151}), (\ref{e198}), (\ref{e199}), and (\ref{e201})
can be combined to give the following complete set of equations that determine the time evolution of the reconnected 
magnetic flux at the resonant surface:
\begin{align}\label{e202}
{\mit\Delta\hat{\Psi}}_s\equiv \tau_s\left(\frac{d\hat{\mit\Psi}_s}{dt} + {\rm i}\,\omega\,\hat{\mit\Psi}_s\right) &= E_{ss}\,\hat{\mit\Psi}_s + E_{sw}\,\hat{\mit\Psi}_w,\\[0.5ex]
\tau_w\,\frac{d\hat{\mit\Psi}_w}{dt} &= E_{sw}\,\hat{\mit\Psi}_s +\tilde{E}_{ww}\,\hat{\mit\Psi}_w
+\hat{I}_c,\label{e203}\\[0.5ex]
\omega(t)& = \omega_{0} - \sum_{p=1,\infty}\left[\alpha_p(t)+\beta_p(t)\right],\label{e210a}\\[0.5ex]
(1+2\,q_s^{\,2})\,\frac{d\alpha_p}{dt} + \left(\frac{1}{\tau_{\theta\,s}}+\frac{j_{1\,p}^{\,2}}{\tau_{M}}\right)\alpha_p
&= \frac{m^{2}\,[y_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,\epsilon_a^{\,2}\,[J_2(j_{1\,p})]^{\,2}}
\,{\rm Im}({\mit\Delta\hat{\Psi}}_s\,\hat{\mit\Psi}_s^{\,\ast}),\label{e205}\\[0.5ex]
\frac{d\beta_p}{dt} + \frac{j_{0\,p}^{\,2}}{\tau_{M}}\,\beta_p
&= \frac{n^{2}\,[z_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,[J_1(j_{0\,p})]^{\,2}}\,
{\rm Im}({\mit\Delta\hat{\Psi}}_s\,\hat{\mit\Psi}_s^{\,\ast}).\label{e206}
\end{align}
Here, 
\begin{align}
\hat{I}_c &= \left(\frac{r_w}{r_c}\right)^m\frac{\mu_0\,I_c}{R_0\,B_z},\\[0.5ex]
\tilde{E}_{ww} &= E_{ww} + \frac{2m\,(r_w/r_c)^{2m}}{1-(r_w/r_c)^{2m}}.\label{e208}
\end{align}
